---
title: "Mini Project 1 Static Maps"
format: pdf
editor: visual
author: Gwynnie Hayes
editor_options: 
  chunk_output_type: console
---

## Importing Data

```{r}
#| warning: false

library(tidyverse)
library(dplyr)
library(sf)
library(tmap)
library(maps)
library(viridis)
library(htmltools)
library(glue)
library(leaflet)

us_states <- map_data("state")

frogs <- read.csv("~/Desktop/15/SDS264/data/most_common_frog.csv") |>
  na.omit()

forests <- read.csv("~/Desktop/15/SDS264/data/forests.csv") |>
  mutate('forest_cover' = forest_area/land_area * 100)

states <- read_sf("https://rstudio.github.io/leaflet/json/us-states.geojson")
```

## Data Source

# Forest

For my numerical maps, I looked at the percentage of land covered in forest in the united states. I found this data in the USDA Forest Service FIA Annual Report, this report is a pdf where I extracted the values that I would need for this data set.

# Frogs

I wanted to look frogs for my categorical data. I found observations of frogs on the website, inaturalist where you can export observations based on specific criteria. I wanted to look at something more interesting then just the most common frog in each state so i looked at the most commonly observed frog that is deemed threatened.

## Joining with Polygons

```{r}
#| warning: false
forests_static <- forests |>
  mutate(state = str_to_lower(state)) |>
  right_join(us_states, by = c("state" = "region")) 
```

```{r}
#| warning: false
frogs_static <- frogs |>
  mutate(state = str_to_lower(state)) |>
  right_join(us_states, by = c("state" = "region"))
```

## Static Maps

```{r}
#| warning: false
forests_static |>
  ggplot(mapping = aes(x = long, y = lat, group = group)) + 
  geom_polygon(aes(fill = forest_cover), color = "black", linewidth = 0.2) + 
  labs(fill = "% Forest Coverage", title = "Percent of Forest Coverage in each US State", x = "Longitude", y = "Latitude", caption = "Data Source: USDA Forest Service FIA Annual Report") +
  coord_map() +
  scale_fill_viridis(option = "viridis", direction = -1) +
  theme_linedraw()
```

alt text: This is a choropleth map of the United States that is looking a the percentage of land area that is covered by forest in each state. The x-axis is the longitude values of the United States which contain values -120 - -80. The y-axis the latitude values of the United States which contain values 25-50. This map is colored by the percentage forest covereage with the most coverage being a dark purple while the least coverage is a bright yellow. From this map we can see that states in the East have a higher percentage of forest coverage while states in the middle have a much lower percent forest coverage, then the west coast has a medium amount of forest coverage.

```{r}
#| warning: false
frogs_static |>
  ggplot(mapping = aes(x = long, y = lat, group = group)) + 
  geom_polygon(aes(fill = frog), color = "black", linewidth = 0.2) + 
  labs(fill = "Frog Common Name", title = "Most Observed Threatened Frog in Each State", x = "Longitude", y = "Latitude", caption = "Data Source: https://www.inaturalist.org/observations?place_id=1&subview=map&taxon_id=25473&threatened") +
  scale_fill_manual(values = c("#800000", 
                               "#9A6324", 
                               "#808000", 
                               "#469990", 
                               "#000075", 
                               "#000000", 
                               "#e6194B", 
                               "#f58231", 
                               "#ffe119", 
                               "#bfef45", 
                               "#3cb44b", 
                               "#42d4f4", 
                               "#4363d8",
                               "#911eb4", 
                               "#f032e6", 
                               "#fabed4", 
                               "#ffd8b1",
                               "#fffac8",
                               "#aaffc3")) +
  coord_map() +
  theme_linedraw()
```
